Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-09-25 paper.tex
restructured report

\usepackage{hyperref} \usepackage{xcolor} \usepackage{ulem}

@@

\begin{document} \title{Pipeline systems and infrastructure for the efficient
and open processing of Big neuroimaging Data} \author{Valerie Hayot-Sasson}
\maketitle \begin{abstract} Text of abstract.  \end{abstract} \tableofcontents
    \chapter{Introduction} intro here -mention hpc, data centre, cloud.
\chapter{BigData and Open science in neuroimaging} \note{Likely will not be
split in two separate sections} \section{large images} \begin{itemize}  \item
        BigBrain \item Micro CT \item EM \end{itemize} \section{large datasets}
        \note{off the top of my head, but need to double check} \begin{itemize}
        \item HCP \item UK Biobank \item ADNI \item OpenFMRI \item BIDS/BIDS
        apps \end{itemize} An effective engine to process such data requires the
        following characteristics \begin{itemize} \item in-memory computing
        \item data locality \item lazy evaluation \end{itemize}
            \section{open-science /data sharing platforms for neuroimaging} is
            open science just data sharing, computing.
            \subsection{Introduction} \subsection{CBRAIN} \subsection{OpenNEURO}
            \subsection{NeuroData} \subsection{Apache Airvata} \chapter{Workflow
            engines for neuroimaging data} \section{Definition of
            workflows/pipelines}

    Broad definition of activities - between the tool and the engine.
    \section{Nipype} \subsection{summary} Nipype\cite{nipype} is a popular
    neuroimaging pipelinening framework written in Python.  It provides users
    with uniform access to the rich ecosystem of neuroimaging software libraries
    (e.g. SPM, FSL, Freesurfer) through its \textit{Interfaces}. \weird{Should
    users want to use their own custom tool, it is also possible to create an
    interface for that}. To execute Python code in Nipype, the \textit{Function}
    interface can be used \tristan{for arbitrary code?}. The framework is also
    easy to use, and thus does not contribute to the steep learning curve new
    researchers must face.

        A pipeline in Nipype, consists of a data analysis. explicit
        \textit{Workflow} of connected \textit{interfaces}. The workflow is a
        directed acyclic graph that connects the outputs of one interface to be
        the inputs of another. A workflow may also consist of connected
        workflows.

                \tristan{The text below is about workflow execution while the
                text above is about workflow description, this should be a new
                paragraph.}  Workflows may be executed in parallel either
                locally or an a cluster through the use of a plugin in the
                Workflow's run function. For debugging, Nipype uses
                \textit{graphviz} to generate a static graph representing the
                Nodes and their relationships. \tristan{debugging might be a
                third theme, in addition to description and execution.
                Provenance might go in debugging.}

         Each \textit{interface} within the workflow must be contained within a
         \textit{Node} or a \textit{MapNode} object, the MapNode class being a
         subclass of the Node class. Wrapping an interface in such objects
         ensures that interfaces are executed within a uniquely named
         directories, which in turn enables provenance tracking. Node objects
         also hash inputs, provide the ability to iterate over inputs and cache
         results. Hash inputs are useful in the case of Workflow recomputation,
         where Nipype will only recompute Nodes whose inputs have changed since
         the previous run.

        Inspired by the MapReduce paradigm, the MapNode copies the interface to
        each input and executes it independently. \tristan{How are inputs
        defined/combined?} A reduce is subsequently performed by the MapNode to
        return the output in the form of a list. This differs from a Node object
        which can only execute the interface on a single input at a time.

        To enable execution of different parameters on the input, Nodes have a
        property known as \textit{Iterables}. Similarly to a MapNode, a copy of
        the \textit{interface} will be made for each input. However, with
        \textit{iterables}, a copy of each dependent node will also be made.

        \subsection{Limitations} Nipype, although meeting the various needs of
        researchers in Neuroinformatics \tristan{vague}, was not designed with
        the processing of Big Data in mind. Moreover, its performance relative
        to current Big Data frameworks remains unknown \tristan{You could refer
        to your MS thesis here}.  \note{nipype performs a lot of disk I/O.
        Recomputation of workflows vs RDD caching}

    \section{PSOM} \subsection{summary} While Python is a popular programming
    language amongst neuroinformaticians, it is not the only one that is
    commonly used. To address the needs of the Matlab/Octave community, PSOM
    (Pipeline System for Octave and Matlab) was developed. Similarly to Nipype,
    PSOM is easy-to-use, can run in distributed environments and can interface
    with tools written in other languages.

        \subsection{pipeline description} A pipeline in PSOM, is a series of
        connected jobs, which can be represented as a (static?) directed acyclic
        graph. The pipeline itself is implemented using the Matlab/Octave data
        type. Each job within the pipeline is assigned a unique name. There is
        only one mandatory field in the job description (command). The remaining
        four optional fields include the list input files (files\_in), the list
        output files (files\_out), the list of files to be deleted after
        execution (files\_clean) and any Octave or Matlab datatype containing
        the remaining parameters necessary for job execution (opt).  Jobs in
        PSOM run in a protected environment where they only have access to the
        parameters that are passed.

        In order to determine whether the opt field was correctly set, the
        generic function \textit{psom\_struct\_defaults} may be applied. This
        function will set the default values, ensure all mandatory fields were
        supplied and issue warnings for all unidentified attributes.

        To define modules in PSOM, a special Octave/Matlab function type, known
        as "bricks" is used. Bricks employ the same parameters as jobs but lack
        the file\_clean parameter. All bricks contain an opt.flag\_test boolean
        flag, which, if set to true, allows the user to perform a dry-run of the
        brick. A dry-run of a brick consists of updating the default parameters
        and validating that the brick can successfully execute with the
        parameters provided. Bricks also have an option to specify the output
        folder (opt.output\_folder) for brick outputs to be stored.

        The pipeline generator is an easy way to create complete pipelines with
        minimal information provided. A pipeline generator may accept at most
        the file\_in and opt arguments. It then builds the pipeline based on
        this information and the default parameters specified in each brick
        applied by the pipeline generator.

        \subsection{pipeline execution}

        To execute a pipeline, the user must provide the pipeline structure and
        the pipeline configuration options. The latter consists of the log
        directory, which must be provided, and the pipepline execution modes
        (e.g., session, background, batch, qsub, msub), in which the default
        mode is \textit{background}. Pipeline execution is achieved through the
        execution of the \textit{psom\_run\_pipeline} command.

        Pipeline execution is broken down into three distinct modules: 1)
        initialization stage, 2) pipeline manager and 3) job manager. The
        initialization stage performs some checks to ensure that the pipeline
        can execute successfully (e.g., verify that pipeline is a DAG, check
        that output files are generated only once, check that all input files
        not generated by the pipeline exist on disk, delete any already existing
        output files), as well as performance checks (e.g.  do not recompute
        pipeline steps that have already been executed, create necessary output
        folders to avoid creating them for each individual job). The
        initialization manager can determine which jobs need to be executed by
        comparing the current pipeline jobs with the details of those previously
        executed stored in the logs folder. The jobs that have been executed can
        have the following states: 1) none, 2) finished and 3) failed. The
        initialization manager stage the jobs with a \textit{none} or
        \textit{failed} status for recomputation. For jobs whose description has
        changed since the previous \textit{finished} execution, the
        initialization will also stage it for recomputation. Should the user
        want to reexecution a previously executed \textit{finished} job, the
        user may force reexecution by setting \textit{opt\_pipe.restart}
        parameter to include all jobs to be reexecuted.

        The pipeline manager is responsible for monitoring the state of the
        pipeline and submitting jobs for execution. Its duties begin after the
        initialization manager has completed execution. Jobs can be submitted by
        the pipeline manager once all their dependencies have been satisfied.
        The dependencies of a job are designated as met if there are none or
        when its parent jobs within the dependency graph are tagged as
        \textit{finished}. In addition, a job is considered to be successfully
        completely when all the dependencies of its children jobs have been
        satisfied. Depending on the pipeline configuration, there may be a limit
        to how many jobs can run in parallel. Once a job has finished executing,
        an empty tag file is created with a filename containing the final state
        in the logs folder. The pipeline manager continuously parses through the
        logs folder in search of these state files to delete and update the
        history file with the job state. Both the pipeline manager and the jobs
        can be executed within the current session, locally in a independent
        session, or in an independent session on a cluster

        Every time a job is submitted by the pipeline manager for execution, a
        job manager is started. It is a Matlab function responsible for creating
        the job profile and generating the job-specific logs and the job exit
        tags. When the job terminates, the job manager will verify that all
        output has been correctly generated and assigns a job tagged based on
        this verification that will be reported back to the pipeline manager.

        \subsection{debugging and provenance}

        The information stored in the logs folder is complete with details
        allowing for both debugging and provenance. The logs folder contains six
        files: 1) PIPE\_history.txt, 2) PIPE\_jobs.mat, 3) PIPE\_status.mat, 4)
        PIPE\_logs.mat, 5) PIPE\_profile.mat and 6) PIPE.mat. PIPE\_history.txt
        includes the execution history of the pipeline manager. PIPE\_jobs.mat
        and PIPE\_profile.mat are both Matlab/Octave files containing each
        executed job as a Matlab variable. They differ in that PIPE\_jobs.mat
        contains the latest executed jobs and can be used to reexecute the
        pipeline at any moment in time and PIPE\_profile.mat contains the
        execution time of each job \note{i think both can be used to reexecute
        pipeline from scratch. Also unclear whether PIPE\_profile.mat contains
        more than just latest jobs}. It is also possible to access any of the
        previously used parameters through PIPE\_jobs.mat. PIPE\_status.mat
        stores the status of each job as a string variable, whereas
        PIPE\_log.mat stores each job's log as a string variable. Job logs, as
        generated by the log manager, consist of a text copy of the execution,
        in addition to the user name, date, time and operating system used to
        execute the job. Finally, all configuration details are stored in
        PIPE.mat.

        \subsection{limitations of PSOM} \section{SPM and matlabbatch}
        \note{This section may not make much sense. Have not really yet fully
        understood matlabbatch} \subsection{summary} SPM is a very popular
        neuroimaging toolkit, and as a result, neuroimaging pipeline engines
        such as Nipype provide interfaces for it. SPM is also equipped with its
        own neuroimaging pipeline engine known as \textit{matlabbatch}.
        Similarly to PSOM and as its name suggest, it is written in Matlab.

        \subsection{pipeline description} Pipelines in matlabbatch consist of a
        series of connected data processing steps known as "modules" \note{DAG??
        probably but not explicitly listed anywhere}. All SPM modules, helper
        modules and their respective dependencies can be defined using SPM's GUI
        \note{can probably also be done through command-line}.

        For any matlabbatch job, it is possible to create a skeleton of it. A
        skeleton is a Matlab script containing a commented list of necessary
        inputs at the top and all code necessary to initialize and run the
        script. Before running, the skeleton will need to be configured with the
        input names and other necessary parameters.

        All configuration options can be set/changed by modifying the
        \textit{cfg\_$\langle type \rangle$ } files.

        Defaults can be set within the configuration files, through a default
        file, or in a \textit{.def} field for leaf nodes \note{terminal
        jobs???}. The latter updates at the start of the new job execution
        whereas the former are configured at the beginning of the
        SPM/matlabbatch startup.

        Virtual outputs represent the dependencies between jobs \note{unsure}.
        They contain three paramaters of which only one is optional: sname,
        src\_output, tgt\_spec. sname is the display name that will be stored in
        the dependecies list, whereas src\_output is a subscript reference used
        to refer to the output that is produced at runtime. The optional
        parameter, tgt\_spec, is a description of the kind of inputs the outputs
        produced can be a dependency for. The virtual output callbacks are
        evaluated when cfg\_(ex)branch, cfg\_choice and cfg\_repeat have the
        correct number of child nodes within all their in-tree nodes.

        \subsection{pipeline execution} Pipeline execution can be achieved
        either through the provided GUI or by the command-line interface. The
        GUI, however, requires significant user interaction, as the user must
        repeat the "Runs" command as many times as necessary to reexecute the
        command on multiple data. The CLI, however, avoids this with the help of
        its \textit{spm\_jobman} interface.

        The first step to running a batch pipeline is to perform the
        initialization. This may include configuring the
        \textit{spm\_defaults.m} code to run in command line and executing the
        \textit{spm(’defaults’, MODALITY)} (where MODALITY represents the
        modality used in the pipeline) and \textit{spm\_jobman(’initcfg’)}.

        After the intialization is complete, a job may be executed using the
        \textit{spm\_jobman(’run’, job[, input1, input2 ...])} command. This
        command requires the job names as well as its necessary inputs specified
        in the correct order. The job can be specified as a variable, the name
        of a script that will generate a job variable or a cell list containing
        a mixture of scripts and job variables.

        \subsection{debugging and provenance} All batch scripts can be saved as
        a Matlab scripts. These scripts not only contain the code that was
        executed, but also all input parameters. Should any script fail in the
        process of executing, an error will be printed to the Matlab console. If
        the user is using the GUI, the error output may also be saved to a
        Matlab file.

        \subsection{limitations of SPM}

    \section{FastR} \subsection{summary} FastR is another Python-based
    neuroimaging pipeline framework that provides added features such including
    access to different versions of the same command-line tool \note{criticizes
    nipype for not having such a feature, but nipype can access boutiques
    tools...so...} and compatibility with various possible filesystems. In
    addition to its added features, FastR also guarantees a simple-to-use
    framework for building pipelines written in any languages, has built-in
    provenance support, provides the seamless ability to parallelize pipelines
    with the use of plugins and avoids the numerical instability issue with the
    simple integration of containerized pipelines. Similary to many other
    pipeline framework, it also employs a DAG to describe data flow.

        \subsection{Pipeline description}

        A pipeline in FastR is a \textit{Network} of connected \textit{Nodes} in
        which the directed edges represent the flow of data.  Each \textit{node}
        contains a JSON or XML template known as a \textit{Tool}.

\textit{Tools} describe the details of the underlying application that is to be
executed at the node. Tools consist of three components: general metadata,
target and interface.  The general metadata provides details on the Tools such
as name, version, authors, etc, whereas the target specifies details on the
execution environment.  The interface, on the other hand, provides details on
the inputs and outputs of the tool.

A JSON schema is required to validate the python object generated by the Tool.
Fastr currently provides two schemas for object validation: the \textit{Tool}
schema and the \textit{FastrInterface} schema. It is, however, possible to add a
custom schema to FastR and define an \textit{Interface} for it.

The \textit{Interface} class is responsible for executing a call to the command
line tool.  It dictates the tags that must be defined within the interface tag
of the \textit{Tool}. For instance, the generic interface class requires details
on the input and output's id, cardinality (number of input values) and DataType
(a python object denoting the details of the Input or Output's data type).
DataTypes can also be grouped, allowing an inputs containing various files types
to be inputted to applications which can support the file types.  Verification
is done by FastR to determine if the linked Input and Output data types are
compatible.

Input samples can consist of one or more values, in which the output is also a
sample consisting of the same number of values. During the instances in which it
is desirable to apply the same task to multiple inputs, a multidimensional array
of samples, known as a collection, can be supplied to a \textit{Node} for
processing. If there are multiple input samples, they can be passed to a node in
pairwise behaviour (e.g. 1-1) or cross-product behaviour (e.g. nxn) \textit{need
to better explain this}.

Output samples may need to be expanded (i.e. a single sample of values becomes
separate samples) or collapsed (i.e. multiple samples become a single sample of
values).  FastR provides data link flow directives for such operations.

In FastR, there are four types of \textit{Nodes}: normal Nodes, Sink Nodes,
Source Nodes and Constant Nodes. Data enters the network through Source or
Constant Nodes and exit through Sink Nodes. Constant nodes differ from Source
Nodes in that their input data are defined by the network \note{Not sure if
saying this correctly}. Normal nodes are responsible for processing the data.
Nodes are connected to each other through links that highlight the dependencies
and form the \textit{Network}. Once the \textit{Network} is defined, it can be
executed.

The data type of the Source Node and Sink Node must be defined in the
\textit{Network} definition.  The data is define at runtime by evaluation of the
uniform resource identifier (URI). Multiple URIs can be defined through the use
of wildcards and searches. Plugins are used to read and write data to different
filesystems specified in the URI scheme. Currently, FastR supports virtual file
system (VSF), csv files and XNAT. Plugins to support other filesystems can
easily be added to FastR. In addition, plugins that can only read, write, or
perform searches can also be created.

        \subsection{Pipeline execution} To commence pipeline execution, the
        \textit{Network}'s execution function is called on the Submission host.
        FastR then analyses the network and its dependencies and breaks down the
        Network into chunks that can be further processed. From the chunks, the
        execution order of the \textit{Nodes} is determined.

During \textit{Node} execution, a job is created for each input/output
combination. Jobs are then executed on the execution host (i.e. local, cluster,
cloud), as determined by the execution engine plugin. FastR currently only
offers support for local and cluster plugins, however can easily be extended to
support other plugins.

While the job is executing, input URIs are translated to their actual
path/values, the execution environment is set up by the \textit{Tool} and the
\textit{Tool} command is executed by the \textit{Interface}. When the
\textit{Interface} returns the results, the output data is validated and
translated to URI format.

After job completion, the results are returned to the Submission Host which
reads the output and updates the network accordingly. Should a chunk be
completed, the next chunk will be processed with the updated information, until
all chunks have been processed and network execution is complete.

        \subsection{Debugging and Provenance} In FastR, a provenance document is
        generated for each resulting file generated in the Prov-N format. Data
        recorded in the provenance documents include \textit{Tool} versions,
        input/output paths/values and their respective checksums, start and end
        time of execution for each \textit{job}, stdout and stderr logs, the end
        status and the complete environment description.

    Similarly to Nipype, the FastR network can be visualized using graphviz.

    \section{LONI} \subsection{Summary} The LONI pipeline processing environment
    facilitates the development, sharing and execution of pipelines. Workflow
    Activities, known as \textit{Modules} in LONI, are described using
    Extensible Markup Language (XML), which can easily shared between user
    groups. LONI runs on the Java platform, allowing it to be platform
    independent and has support DRMAA clusters and symmetric multiprocessing
    (SMP) systems. In order to capture provenance information, LONI is also
    equipped with a provenance manager.

        \subsection{Pipeline description} LONI uses the dataflow programming
        model to represent the pipeline. Data flows between \textit{Modules}
        through \textit{data pipes}. Sub-pipelines are known and
        \textit{Pipelets}. Modules are defined by the input arguments/streams
        and output arguments/stream. Arguments in LONI may be specific
        filetypes, flags or values listed in the order in which the application
        requires. The provided arguments may be required or optional and can
        also be duplicated, if necessary. File inputs in the form of URLs are
        accepted by LONI and will be downloaded at runtime. List of inputs may
        also be provided to process multiple inputs with copies of the same
        module in parallel (batch processing). Conversely, it is also possible
        to group list of inputs so as to process them all by the single module
        \note{Need to express this better}. Input/Output arguments may be added
        to the table of global variables for easy access throughout the
        pipeline. It is possible to apply variable modifiers to the variables
        stored within the table.

Intermediary outputs as well as the file output are treated as temporary by
LONI. They are scheduled for deletion before a new pipeline is executed. It is,
however, possible to preserve these files.

        \subsection{Pipeline execution} As a module completes it execution, its
        output gets transmitted through the dataflow and the dependent modules
        are signaled that their input data has arrived.  Modules who have met
        all their dependency requirements are ready for execution.  Should a
        module have no dependecies, it can be executed immediately. Modules
        ready to be executed are placed in a FIFO launch queue to avoid
        straining the platform.

       LONI offers a built in client/server strategy that uses Java's remote
       method invocation (RMI) libraries to communicate between servers. This
       enables modules, pipelets or pipelines to be executed on servers
       different from the client. It also enables full pipelines to consist of
       connected modules and pipelets located on different servers. It is
       favorable to have such a configuration in LONI, as the tools required for
       processing may not be available on client server, but located in another.
       The client is responsible for dictating all file transfers and module
       launches that are to occur. The client authenticates with the servers
       through a one-way hash. Only files produced by the pipeline are
       accessible to the client and only temporary files created by the server
       may be written to the server.

        \subsection{Debugging and provenance} For each dataset processed in a
        LONI pipeline, there exists a corresponding provenance (.prov) XML file.
        The \textit{Provenance Editor}, a standalone platform-independent
        application, reads the metadata from the image headers, extracts
        relevant provenance information from them and creates the provenance
        files.  This data can be modified by the user in the to correct
        inaccuracies or append additional details before being saved. During
        pipeline execution, the provenance file will be edited by the LONI
        pipeline to include details on each processing step, the environment in
        which the data was processed and tool binaries used.

    \section{Automatic Analysis} \subsection{Summary} Automatic analysis (aa) is
    another Matlab-based modular workflow engine, having started of as a
    workflow engine for SPM. User may define their pipeline through a Matlab
    user script and specify the tasks in addition to their input and output
    types through an XML module interface. Modules used in pipelines must be
    integrated into aa which can be easily accomplished through XML interfaces.
    aa workflows can be executed locally, or in parallel through qsub or condor.

        \subsection{Pipeline description} aa pipelines have two main components:
        1) the Matlab user script and 2) the XML tasklist. The user script is a
        script that defines which data will be processed, any necessary
        parameter, and how it will be processed. aa provides an example user
        script, which can be modified to meet user's needs. The entire pipeline
        in aa is defined in a single structure, known as aap, which makes the
        task of record-keeping simple.

        The XML tasklist is an ordered list of all the modules that will be
        applied to the data, that is read by the user script. This list does not
        specify any dependencies as those dependendecies can be inferred by the
        pipeline through analysis of the module interfaces. Modules can be
        separated into two sections: initialization and main. Modules listed in
        the initialization section are to be executed for every task.
        Conversely, those modules liste in the main section are only executed
        once for every task, unless otherwise specified.

        In order to ensure that the correct output data is being operated on,
        each task is provided with its own unique folder. The aa engine will
        transmit the necessary input data to the folder before task execution
        and also acquiring the outputs after task completion. Output folders are
        numbered such that if a given module is invoked multiple times, each of
        those folders are assigned an incremental ordering corresponding to the
        order in which they have been executed \note{sentence might not make
        sense}. Should all data be required for each task in the tasklist, all
        the data will be copied to each task folder.

    Similarly to a pipeline, each module is comprised of an XML interface and a
    Matlab source file. A module's XML file consist of specifications on
    required input and output modalities, as well as the desired domain. One of
    the most important properties include the domain at which the module
    operates. While there exist mainy domains for modules and new ones can
    easily be created and integrated by the user, the most prominently used ones
    include: "study", "subject", "session". As their name suggests, a module
    with the domain "study" will only be executed once for each time it appears
    in the tasklist, whereas "subject" executes the module for each subject
    within the task and "session" executes once for each session within a
    subject.

    In addition to module domains, another important module property is the
    input and output data stream types. Data stream types denote what input and
    output data type each module produces. Commonly used data stream types
    include "epi", "structural" and "dicom\_header".

    Optional fields specified in a module include qsub, for estimates of
    requirements for parallel processing, and \textit{performanceoutput}, for
    garbage collection.

   Analysis parameters can easily be configured in aa. Any variable defined by
   the \textit{aaprecipe} command may be modified. Parameters can also be
   defined for an instance of a module within the XML tasklist. Alternatively,
   it is also possible to create an XML file that inherits parameters from
   another XML file and overrides some or all of the parameters.

   To set defaults and site-specific configurations, it is possible to either
   create a local config file or by creating another parameter default file
   which inherits the original one.

    aa also supports the creation of branched tasklist. This allows a section of
    the pipeline steps to undergo different tasks or the same tasks in a
    different order in one pipeline rather than having to create separate
    pipelines.

        \subsection{Pipeline execution} Once the definition of the user-script,
        XML tasklist and modules is complete, the pipeline is ready for
        execution. Dependencies in aa are established based on input and output
        data stream types. The aa engine attempts to match the last task
        executed with a corresponding data stream type to the current type. It
        is, however, possible to specify which module provides the input to the
        current module by using fully qualified stream references.

        During execution, a map of all the modules that need to be executed and
        their respective dependencies is generated. Modules that have completed
        contain a flag file indicating their completion within their instance's
        root directory. Such modules will not be re-executed by the engine.
        Additionally, if it is required that a previous module is re-executed,
        the done flag is removed from all dependent modules.

        Should part of the processing be performed on a different server it is
        possible to connect the pipelines together through an analysis script.
        The aa engine will be reponsible for matching the data streams between
        the servers in order to avoid the copying of data and modules.

        \subsection{Debugging and provenance} In order to retrace the steps of
        previous executions, aa keeps a detailed record of the events that
        occurred. These detailes are stored in a Matlab object \note{similarly
        to PSOM}, which permits reexcution of a previous run or extraction of
        specific parameter values.

        Additionally, to verify the correctness of the results produced, aa
        implements tools for quality control and comes with a dedicated module
        for low-level quality control.

    \section{Pydiper} \subsection{Summary}

        Pydiper is a neuroimaging registration workflow engine implemented in
        Python. This differs from other workflow engines in that it is specific
        to the process of image registration, but can, however, be extended for
        use with any analysis process. Its main features include being easy to
        use, particularly for non-developpers, undertaking a modular approach,
        giving the ability to access output files at any stage, and eliminating
        duplicate stages. Pydiper also provides the added advantage of being
        equipped with four commonly used registration pipelines, such that
        unexperienced researchers can use such pipelines out-of-the-box.

        \subsection{Pipeline description} Pydiper workflows can be broken down
        into three components: atoms, modules and applications. An application
        is the command line application that encapsulates the pipeline which
        consists of modules and atoms. Atoms are wrappers for individual
        operations, typically frquently used MINC tools. Modules are reponsible
        for linking atoms together.

        Pydpiper provides three main classes to use to develop pipelines. These
        include the \textit{PipelineStage}, \textit{CmdStage} and
        \textit{Pipeline} classes. The \textit{PipelineStage} is the base class
        wrapper for all pipeline tools.  Pipeline tools can be written in Python
        or can be a command-line application.  For the tools that are
        command-line applications, the \textit{CmdStage} class, a child class of
        \textit{PipelineStage}, must be used. Input to \textit{CmdStage} is in a
        adequately parsed array format containing the specific command-line tool
        and its arguments. Reusable subclasses of the \textit{CmdStage} class
        can be created, or \textit{CmdStage} can be called directly.

        To combine pipelines, the \textit{AddPipeline} function of the
        \textit{Pipeline} class may be used.

        The base class common to all Pydpiper applications is
        \textit{AbstractApplication}.  Classes that inherit from
        \textit{AbstractApplication} are command-line tools, with their options
        set by the \textit{AbstractApplication} class. The class is also
        responsible for initializing or resuming the pipeline, setting up a
        logger, and launching the pipeline daemon.

        Pydpiper atoms inherit from the \textit{CmdStage} class in order to wrap
        the MINC Tools.  They require as input a MINC file, and may additionally
        also require a target MINC file as second argument. These arguments can
        be in the form of a string filename or a file handler, however, the same
        format must be used for the two. An atom may also accepts various
        optional arguments. Using the inputs provided, the atom will put
        together the command to be executed.

        \subsection{Pipeline execution}

        In Pydpiper, there are two main classes,\textit{Pipeline} and
        \textit{PipelineExecutor}, that monitor and manage pipeline execution.
        The \textit{Pipeline} class's main responsibility is to infer
        dependencies between pipeline stages from their inputs and outputs. It
        also manages a queue of runnable stages (stages in which all their
        dependencies have been met), ensuring that any runnable stage is placed
        in the queue. It also keeps track of the status of the runnable stages
        (running, finished, or failed). Should an already existing stage be
        added to the pipeline, it is automatically skipped by the
        \textit{Pipeline} class.

        As stages may be added to incomplete pipelines, Pydpiper provides the
        functionality of resuming pipelines. This is achieved through a pickling
        mechanism implemented by the \textit{Pipeline} class. The pickled
        components include the directed acyclic graph, the array of pipeline
        stages, the stage counter, the hashes of all stages and each of their
        output files, and the status of each stage.

        The \textit{PipelineExecutor} class, on the other hand, is responsible
        for the efficient processing of the pipeline through parallelization.
        Pydpiper's architecture mimics that of client/server, in which the
        server is the pipeline daemon started by the command-line executable and
        the executors undertaking the role of the clients. The pipeline's
        executors each consist of a thread pool in which each thread is
        responsible for running a single pipeline stage from the queue of
        runnable stages. To obtain a stage for execution, the executor polls the
        server to obtain the next available stage. Should the executor have
        enough memory and processors to execute the stage, it will be executed
        by the thread and release memory and processors and repoll the server
        once stage has completed.  Otherwise, the executors will sleep for
        certain duration and repoll the server again.  Should there be any
        failed stages, the pipeline will be shut down after all possible stages
        have finished executing. If additional executors are required by the
        pipeline, they may be added at any time through the command-line.
        Details related to the number of threads, number of threads per executor
        and the amount of memory can be specified through the command-line
        options.

        \subsection{Debugging and provenance}

        As mentioned previously, Pydpiper uses a pickling mechanism to gather
        important pipeline details during execution in order to allow for the
        pipeline to be resumed should any stages have failed.  In addition, to
        keep track pipeline stage interdependencies in addition to file
        versioning, Pydpiper utilizes the helper class
        \textit{RegistrationPipeFH} and its parent \textit{RegistrationFHBase}
        to create file handlers of the files.

        \note{from github wiki} For debugging, a Pydpiper pipeline creats its
        own log file and executors typically also each create their own log file
        that are stored in a \textit{logs/} subdirectory. Verbosity of logging
        information can be managed through the setting of the PYRO\_LOGLEVEL
        environment variable. In addition, individual stage stdout/stderr is
        also redirected to a log file.

        To verify the status of an executing pipeline, the
        check\_pipeline\_status.py may be executed.

\chapter{Other domains and general scientific worflow engines} \section{Taverna}
Taverna is an open-source bioinformatics workflow engine, designed to allow
users to combine distributed web services, other service types (e.g. BioMart
queries or R-scripts) and local scripts into complex analysis pipelines. The
main goal of Taverna is to give a unified access to the plethora of services
already existing on the web without requiring the user to download the services
locally or to have knowledge on all of the platforms and programming languages
the services are hosted under.

        \subsection{Pipeline description} Workflows in Taverna are built by
        combining web services, local scripts and other services together. Web
        services compatible with Taverna must comply to the Web Service
        Description Language (WSDL) using the Simple Object Access Protocol
        (SOAP protocol).  These services can be added to the BioCatalogue or the
        Biodiveristy catalogue for sharing and reuse of the web services.

        Workflows can be built in the Taverna workbench. Taverna, through the
        help of the BioCatalogue and the Biodiversity Catalogue is able to
        provide the user with a variety of webservices to combine into workflows
        in a drag-and-drop fashion. Workflows may also be composed of
        sub-workflows. As service inputs and outputs may not match between
        services, shims are provided.  Similarly to the services, the
        user-generated workflows can be uploaded the myExperiment repository to
        be shared with other user.

        \subsection{Pipeline execution} The workflows can be executed remotely
        utilizing the Taverna Server. The Taverna Server is a an application
        that provides the full library of possible workflows through
        webservices. It has an interface, known as the Taverna Player, that
        enables workflow execution to occur through a web-interface or through
        third-party clients. The Taverna Server, however, is limited in that it
        can only execute workflows available on the Server. Taverna Lite is an
        alternative to the Taverna Server which allows users to upload and
        execute workflows from myExperiment or elsewhere.

        \subsection{Debugging and provenance} Provenance capture is achieved
        through the Taverna Provenance Suite. Taverna does not capture
        provenance of the workflow definition, but generates a new UUID after
        each definition change. Workflow run provenance captures information on
        "individual processor iterations" and their respective inputs and
        outputs and stores this information in a database. The results can then
        be exported as a workflow trace or as a PROV-O RDF graph using the
        Taverna-PROV Plugin. The model format used for exported provenance uses
        both the W3C PROV model and Open Provenance Model format.

    \section{Pegasus} Pegasus is a general-purpose scientific workflow engine
    that specializes in mapping abstract descriptions of workflows to resources
    for efficient processing.  Similarly to other workflows, a Pegasus workflow
    can be represented by a DAG. This DAG is described by the user in DAX format
    (Directed Acyclic graph in XML).  As the DAX abstract workflow is
    resource-independent, Pegasus can convert the abstract workflow into an
    execution workflow that is catered to the resources available.

        \subsection{Pipeline description} Within the DAX, the nodes are the
        tasks and the execution order is represented by the edges.  In the
        instance of nested graphs, the nodes in the DAX may also represent a
        DAX. Pipeline inputs and outputs, as well as pipeline executables, are
        represented by logical identifiers.

>>>>

\usepackage{titlesec}
\titleformat{\chapter}
  {\normalfont\LARGE\bfseries}{\thechapter}{1em}{}
\titlespacing*{\chapter}{0pt}{3.5ex plus 1ex minus .2ex}{2.3ex plus .2ex}
\usepackage[left=2cm, right=2cm, top=2cm, bottom=2cm]{geometry}
\usepackage{hyperref} 
\usepackage{xcolor} 
\usepackage{ulem}

@@

\begin{document} 
    \title{Pipeline systems and infrastructure for the efficient
            and open processing of Big neuroimaging Data} 
    \author{Valerie Hayot-Sasson}
    \maketitle

    \begin{abstract} Text of abstract.  
    \end{abstract} 
    \tableofcontents
    \chapter{Introduction}
    In recent years, the volume of neuroimaging data acquired has exceeded
    both the storage and computation capacity of a standard research 
    lab workstation. With the advancement of data sharing technologies, this 
    data has been made widely available, with the only limiting factor impeding
    research on this data being access to infrastructure and software that 
    enables efficient processing on such infrastructure. Many research labs do
    have access to high performance computing (HPC) clusters, however, without
    efficient software to process this data, processing can take an excessive 
    amount of time. Ensuring efficient processing of data is complex and 
    likely beyond the knowledge of an average neuroimaging researcher. As a 
    result, frameworks for the efficient processing of neuroimaging data need
    to be developed.

    Neuroimaging big data comes in two formats: 1) large images and 2) large 
    datasets. Large images consist of a singular images ranging from 100s of
    gigabytes to terabytes in size. A well-known example of such an image is the
    BigBrain \note{cite here}, a histological image of a 69 year-old man. 
    At its highest resolution, it is 1x1x20micrometers \note{fix this} or 1TB in
    size. Other examples of large images can be found in electron microscopy 
    (EM), polarized light images (PLI) and micro coherence tomography (microCT)
    \note{need to do more research here}. Images at such high resolution are
    important to researchers as they give insights into aspects not otherwise
    detectible in images at lower resolutions. However, due to their size and 
    lack of resources available to process such images, research on this data 
    remains limited.

    In contrast, large datasets are consists of many images, typically arising
    from multiple different subjects, that are too big to fit in storage. 
    The images alone being small enough for an average researcher to process. 
    Example of these datasets include the UK Biobank, HCP Project, ADNI dataset,
    OpenfMRI, amongst many others. These datasets are very large. For instance, 
    the UK Biobank is expected to exceed \note{X} Petabytes in size. As subsets
    of these datasets are manageable for the average researcher to process, 
    much analysis has gone into them. Moreover, unlike for large images, 
    toolkits have been developed for researchers to process this data. However,
    the toolkits do not enable researchers to process these entire datasets as
    a whole. Therefore, it is crucial to develop efficient frameworks for the 
    processing of these datasets in order for researchers to harness the data
    in its entirety.

        \section{Platforms for  neuroscience}
        \section{Infrastructure}

    \chapter{Workflow composition}
    \chapter{Performance}
    \chapter{Reproducibility}
    \chapter{Portability}
    \addcontentsline{toc}
        {chapter}{Bibliography} 
        \bibliography{bibliography}
        \bibliographystyle{ieeetr}
\end{document}
